Deep Neural Architectures for Discourse Segmentation in E-Mail Based Behavioral Interventions

### 2018-11-14 paper-amia-summit-2019/manuscript.tex
added new table for dataset

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors. Each session represents an MI intervention delivered via email. We consider segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not. In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark. Each sample is classified into either a ``new segment'' or a ``same segment'' based on whether there should be a segment break after the current word or not. In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

@@

\textbf{Convolutional Recurrent Neural Network}: CRNN is a deep neural network architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, which consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output. 
E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = [n_w;n_p]$ of size 310. The primary purpose of convolution is to extract new features depending on the neighboring words. We used 1D convolution. In this layer, one filter is responsible for extraction of one feature. After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ useful features are produced in the convolution layer for each word. A max pooling operation was then performed over time to find the most significant features in a textual fragment. The bidirectional recurrent layer receives new features extracted in the convolution layer. RNN is usually used to capture long-range dependencies in a sequence of observations. Moreover, bidirectional RNNs are capable of capturing both past and future contexts through forward and backward traversals of a sequence. The purpose of the fully connected layer is to use the output of bidirectional RNN layer for classifying each word or punctuation into ``new segment'' and ``same segment'' classes. Since a fully connected layer has a greater number of parameters, they are more likely to excessively co-adapt with these parameters, which may cause over-fitting. To prevent this, we utilized dropout by randomly disregarding 50\% of the connections to the fully connected layer. After that, logistic sigmoid is applied to the output layer to provide the probability of whether the word or punctuation precedes a new segment or not. We experimentally determined the optimal parameters and found that the best performance is achieved with 5-fold cross-validation when filter length in convolutional layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as a convolutional layer activation function, RNN layer activation is hyperbolic tangent and the number of recurrent units in RNN is 200. Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{source code will be available at https://github.com/teanalab/xxxxx upon acceptance of this paper for publication}

>>>>

\section*{Methods}
\subsection*{\textit{Data collection}}
The experimental dataset for this work was constructed from 49 e-Coaching sessions, which include a total of 3,138 segmented and annotated MI behaviors \hl{shown in Table}~\ref{tab:datastat}. Each session represents an MI intervention delivered via email. We consider segmentation as a binary classification problem, in which each word or punctuation mark is annotated with a label of 1 or 0 to indicate whether it precedes a new segment or not. In total, the dataset consists of 95,777 words and 7,140 punctuation marks and includes 3,138 ``new segment'' and 99,779 ``same segment'' instances. In this study, we experimented with traditional machine learning methods, such as Conditional Random Field (CRF)\cite{lafferty2001conditional} and deep learning methods, such as Multi-Layer Perceptron (MLP),\cite{rumelhart1986learning} Bidirectional Recurrent Neural Network (bidirectional RNN or BRNN)\cite{schuster1997bidirectional} and Convolutional Recurrent Neural Network (CRNN).\cite{treviso2017sentence} For an MLP model, training and testing instances were created based on a sliding window of $2n$ words or punctuation marks over each position in a given input sequence, such that each instance consists of a pair of the next $n$ words or punctuation marks and prior $n$ words or punctuation marks, including the current word or punctuation mark. Each sample is classified into either a ``new segment'' or a ``same segment'' based on whether there should be a segment break after the current word or not. In the case of CRF, BRNN and CRNN models, an email was taken as an input sequence, POS tags and word embeddings of each word or punctuation mark were used as input and binary labels (1 or 0) corresponding to ``new segment'' and ``same segment'' classification decisions were considered as the model output of a model. In the gold standard, words or punctuations within the same segment were assigned the label of 0 and the last word or punctuation mark of a segment were assigned the label of 1.

\begin{table}[ht]
\centering
\caption{\textbf{\hl{Summary of the experimental dataset}}}
\label{tab:datastat}
 \begin{tabular}{|l|l|l|l|l|l|l|l|l|l|}
  \hline
   \multirow{2}{*}{\textbf{Sessions}} & \multirow{2}{*}{\textbf{Instances}} & \multicolumn{2}{|c|}{\textbf{Class Segments}} & \multicolumn{2}{|c|}{\textbf{Tokens}} & \multicolumn{2}{|c|}{\textbf{Emails}} & \multicolumn{2}{|c|}{\textbf{Annotation}} \\\cline{3-10}
   &  & \textbf{new}  & \textbf{same} & \textbf{word} & \textbf{punctuation}  & \textbf{patient} & \textbf{provider} & \textbf{method}  & \textbf{codes} \\ \hline    
 49 & 102,917 & 3,138 & 99,779 & 95,777 & 7,140 & 281 & 330 & MYSCOPE & 115 \\ \hline
  \end{tabular}
\end{table}

@@

\textbf{Convolutional Recurrent Neural Network}: CRNN is a deep neural network architecture,\cite{treviso2017sentence} shown in Figure~\ref{fig:crnn}, which consists of 5 layers: 1) input layer 2) embedding layer 3) convolution layer with max pooling 4) BRNN layer 5) fully connected layer with dropout and sigmoid output. 
E-coaching email exchanges are represented as a sequence of $m$ words and punctuations, which are fed into the input layer to produce a $m \times n_e$ matrix after fetching the pre-trained word vectors. When POS tags are used in combination with word embeddings, 10-dimensional POS vectors are concatenated with 300-dimensional word vectors to obtain new vectors $n_e = [n_w;n_p]$ of size 310. The primary purpose of convolution is to extract new features depending on the neighboring words. We used 1D convolution. In this layer, one filter is responsible for extraction of one feature. After applying $n_f$ different filters with zero-padding on both sides of the input text, $n_f$ useful features are produced in the convolution layer for each word. A max pooling operation was then performed over time to find the most significant features in a textual fragment. The bidirectional recurrent layer receives new features extracted in the convolution layer. RNN is usually used to capture long-range dependencies in a sequence of observations. Moreover, bidirectional RNNs are capable of capturing both past and future contexts through forward and backward traversals of a sequence. The purpose of the fully connected layer is to use the output of bidirectional RNN layer for classifying each word or punctuation into ``new segment'' and ``same segment'' classes. Since a fully connected layer has a greater number of parameters, they are more likely to excessively co-adapt with these parameters, which may cause over-fitting. To prevent this, we utilized dropout by randomly disregarding 50\% of the connections to the fully connected layer. After that, logistic sigmoid is applied to the output layer to provide the probability of whether the word or punctuation precedes a new segment or not. We experimentally determined the optimal parameters and found that the best performance is achieved with 5-fold cross-validation when filter length in convolutional layer is 7, number of filters is 100, max-pool size is 3, ReLU is used as a convolutional layer activation function, RNN layer activation is hyperbolic tangent and the number of recurrent units in RNN is 200. Adam\cite{kingma2014adam} with 50 epochs, the batch size of 32 and learning rate of 0.001 was used for optimization and the early stopping strategy was applied.\footnote{\hl{https://github.com/teanalab/eCoaching-Text-Segmentation}}
